---
title: "Census Population Estimates"
author: "Christopher Walker"
date: "October 16, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## County Population Estimates

This is based on a US Census Bureau dataset available at:

https://www.census.gov/data/datasets/2017/demo/popest/total-cities-and-towns.html

```{r read}
all_data <- read.csv("sub-est2017_all.csv", stringsAsFactors = FALSE)
```

The population estimates can mapped by county using the ggplot2, maps, and mapdata libraries. This works for most counties.

```{r county_mapping}
suppressMessages(library(dplyr))
library(maps)
library(mapdata)
library(ggplot2)
county_map <- map_data("county")
county_data <- all_data %>% filter(SUMLEV == 50) %>%
  mutate(subregion = gsub(" county", "", tolower(NAME))) %>%
  select(subregion, POPESTIMATE2017)
county_map <- inner_join(county_map, county_data, by="subregion")
ggplot(data = county_map,
       aes(x = long, y = lat, group = group)) +
  coord_fixed(1.3) +
  geom_polygon(aes(fill = POPESTIMATE2017)) +
  scale_fill_gradient(trans = "log10") +
  theme_void()
```

## State Population Estimates

The population estimates can be mapped by state as well.

```{r state_mapping}
state_map <- map_data("state")
state_data <- all_data %>% filter(SUMLEV == 40) %>%
  mutate(region = tolower(NAME)) %>%
  select(region, POPESTIMATE2017)
state_map <- inner_join(state_map, state_data, by="region")
ggplot(data = state_map,
       aes(x = long, y = lat, group = group)) +
  coord_fixed(1.3) +
  geom_polygon(aes(fill = POPESTIMATE2017)) +
  scale_fill_gradient(trans = "log10") +
  theme_void()
```

## Plotting all years

The population estimates can be viewed year over year, although this turns out to not be very interesting for most cases. It is too short of a time span to see noticeable population shifts at the county or state level in these.

```{r all_years}
library(tidyr)
all_years <- all_data %>% gather(POPESTIMATE2010,
                                 POPESTIMATE2011,
                                 POPESTIMATE2012,
                                 POPESTIMATE2013,
                                 POPESTIMATE2014,
                                 POPESTIMATE2015,
                                 POPESTIMATE2016,
                                 POPESTIMATE2017, key = "year", value = "popestimate") %>%
  mutate(year = as.numeric(substr(year,12,16)))
colorado_map <- map_data("county") %>% filter(region %in% c("colorado"))
colorado_data <- all_years %>% filter(SUMLEV == 50, STNAME %in% c("Colorado")) %>%
  mutate(subregion = gsub(" county", "", tolower(NAME))) %>%
  select(subregion, popestimate, year)
colorado_map <- inner_join(colorado_map, colorado_data, by="subregion")
ggplot(data = colorado_map,
       aes(x = long, y = lat, group = group)) +
  coord_fixed(1.3) +
  geom_polygon(aes(fill = popestimate)) +
  theme_void() +
  facet_wrap(vars(year)) +
  ggtitle("Colorado Population Estimates Over Time")
```

```{r city}
Colorado <- all_years %>%
  filter( SUMLEV == 40, STNAME == "Colorado") %>% select(year, popestimate)
ggplot(Colorado, aes(x = year, y = popestimate)) +
  geom_histogram(stat = "identity") +
  ggtitle("Colorado Population Estimate Over Time")
```